Spanner: Google's Globally-Distributed Database \newline Seminar Report

### 2016-02-16 paper/paper.tex
Iterate Discussion

\section{Discussion \tbr}
\label{sec:discussion}

The system introduced by Google is definitely impressive because they were able
to establish a notion of global wall clock time (that's how they call it in their
presentation) which enabled them to guarantee serialization order for
globally-distributed transactions. To the best of the author's knowledge, Spanner
was the first system to provide much stronger consistency and time semantics while
data is sharded and replicated at global scale. Google also proved all
'\emph{Relational databases are dead}' people wrong even though systems like
Bigtable helped creating the NoSQL trend. Reading about a system like Spanner is
definitely more than a vital sign of relational database systems. From a
developer's point of view, this is interesting because Spanner provides
the more intuitive schematized view of tables. In contrast to NOSQL database
systems, Spanner could be used as replacement for other relational database
systems, e.g. MySQL, directly. Spanner is also highly flexible to schema changes
because of the atomic schema change transactions.

However, I believe that rebuilding a Spanner-like system is difficult for
small- to mid-sized companies because there is a lot of hardware involved, for
example all the GPS and atomic clocks. Unless this hardware infrastructure is
not available it is also hard to reimplement the TrueTime API. Nonetheless,
companies which need a globally-distributed database with all the guarantees
Spanner provides are likely to be almost as big as Google. Hence, such companies
may also be able to build a Spanner-like system. If every datacenter would provide
GPS and Armageddon time masters independently, the reimplementation of such a
system would be much easier.

Another part that may influence the performance of Spanner is the distributed
file system, i.e. Colossus, the successor of the Google File System (GFS). I did
not find reliable information about the implementation or the performance of
Colossus. It is unknown how much impact this distributed file system has on the
performance of Spanner. Thus, it is hard to say if a reimplementation of Spanner
would be able to compete with Spanner's performance.

Of course, such a system may also be built upon e.g. the Network Time Protocol
(NTP). But this would result in poor performance compared to Spanner because the
error interval would be much larger compared to $2\cdot\epsilon$. And besides
providing distributed general-purpose transactions, the performance of Spanner
makes the whole system that impressive. A system using e.g. NTP would never be
able to achieve this performance while guaranteeing serialization order. Spanner
can synchronize multiple systems without any communication between these systems.
NTP-based systems would have to communicate with each other. For systems far apart
such a system would never be able to compete with Spanner.

Spanner is robust to GPS blackouts because of the usage ofatomic clocks, even
though the performance of Spanner may suffer due to high values of $\epsilon$.
Would be interesting to see how such a blackout affects Spanner's performance.

It would also be interesting how failures of the two singletons per universe,
i.e. universemaster and placement driver, affect Spanner. Since the universemaster
is just a debugging console, most probably a failure in this component will not
affect Spanner at all. This would probably be a totally different story if the
placement driver does not operate correctly. First of all, there would be no
load balancing between zones. Hence, zones could be run out of space if the
placement driver is down long enough. Furthermore, updated replication
constraints, would not be propagated since the placement driver is responsible
to move the data if the replication constraints are modified.

In terms of applications, applications which read a lot but write rarely or little
data would benefit the most from a system like Spanner. If an application uses
lots of replicas which are spread across the globe and writes large amounts of data
regularly, Spanner may not provide the performance mentioned in the original
paper. However, Spanner may still outperform both, Megastore and Bigtable while
providing much stronger consistency guarantees.

>>>>

\section{Discussion}
\label{sec:discussion}

The system introduced by Google is definitely impressive because they were able
to establish a notion of global wall clock time (that's how they call it in their
presentation). Global wall clock time enables them to guarantee serialization order
for globally-distributed transactions. To the best of the author's knowledge, Spanner
was the first system to provide much stronger consistency and time semantics while
data is sharded and replicated at global scale. Google also proved all
'\emph{Relational databases are dead}' people wrong even though systems like
Bigtable helped creating the NoSQL trend. Reading about a system like Spanner is
definitely more than a vital sign of relational database systems. From a
developer's point of view, this is interesting because Spanner provides
the more intuitive schematized view of tables. In contrast to NoSQL database
systems, Spanner can be used as replacement for other relational database
systems, e.g. MySQL, directly. Spanner is also highly flexible to schema changes
because it supports atomic schema change transactions.

However, I believe that rebuilding a Spanner-like system is difficult for
small- to mid-sized companies because there is a lot of hardware involved, for
example all the GPS and atomic clocks. Unless this hardware infrastructure is
available it is hard to reimplement the TrueTime API. Nonetheless,
companies which need a globally-distributed database with all the guarantees
Spanner provides are likely to be almost as big as Google. Hence, such companies
would also have the resources to build a Spanner-like system. An Open-Source
reimplementation would only be possible if every datacenter would provide GPS
and Armageddon time masters on their own.

A component that may influence the performance of Spanner is the distributed
file system, i.e. Colossus, the successor of the Google File System. The internet 
does not provide reliable information about the implementation or the performance
of Colossus. It is unknown how much impact this distributed file system has on the
performance of Spanner. Thus, it is hard to say if a reimplementation of Spanner
would be able to compete with Spanner's performance if a different distributed
file system is used.

Of course, a distributed database system may also be built upon other concepts,
e.g. the Network Time Protocol (NTP). But this would result in poor performance
compared to Spanner because the error interval would be much larger then the one
used by Spanner, i.e. $2\cdot\epsilon$. And besides providing distributed
general-purpose transactions, it is the performance of Spanner that impressives
the most. A system using NTP would never be able to achieve this performance
while guaranteeing serialization order of transactions. Spanner can synchronize
multiple systems without any communication between these systems. NTP-based
systems would have to communicate with each other. For systems far apart such a
NTP-based solution would never be able to compete with Spanner.

Spanner is robust to GPS blackouts because of the usage of atomic clocks, even
though the performance of Spanner may suffer due to high values of $\epsilon$.
Would be interesting to see how a GPS blackout would affect Spanner's performance.

It would also be interesting to know how failures of the two singletons per
universe, i.e. universemaster and placement driver, would affect Spanner. Since
the universemaster is just a debugging console, most probably a failure in this
component will not affect Spanner at all. This would probably be a totally
different story if the placement driver does not operate correctly. First of all,
there would be no load balancing between zones. Hence, zones could run out of
space if the placement driver is down long enough. Furthermore, updated
replication constraints, would not be propagated since the placement driver is
responsible to move the data if the replication constraints are modified.
Therefore I do not understand why there is no redundancy in placement driver.

In terms of applications, applications which read a lot but write rarely or little
data would benefit the most from a system like Spanner. If an application uses
lots of replicas which are spread across the globe and writes large amounts of data
regularly, Spanner may not provide the performance presented in the original
publication. However, Spanner may still outperform all other distributed database
systems while providing much stronger consistency guarantees.
